---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
```

# Downloading and wrangling CMIP6 model output for the Biological Dynamics of Forest Fragments Project region

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/426372968.svg)](https://zenodo.org/badge/latestdoi/426372968)

<!-- badges: end -->

The goal of this repository is to download, crop, and spatially aggregate historical and projected climate data from CMIP6 models for an area centered around the Biological Dynamics of Forest Fragments Project (BDFFP).

# Data sources

Original data were downloaded from two sources: The Copernicus data portal (https://cds.climate.copernicus.eu/cdsapp#!/dataset/projections-cmip6) that provides the ability to spatially crop before downloading .nc files and ESGF (https://esgf-node.llnl.gov/search/cmip6/) for combinations of source, experiment, and variable that were not available from Copernicus.  See .csv files in the `metadata/` directory for a list of files that were not available from Copernicus and were consequently downloaded directly from ESGF.  There were a few that didn't download using the scripts in `R/` and I had to download manually and then crop using the `R/manual_crop.R`.

Observed historical data used for validation of CMIP6 models was provided by [Xavier et al. (2016)](https://doi.org/10.1002/joc.4518). The raw data files are accessed through http://careyking.com/data-downloads/ (direct link: https://utexas.app.box.com/v/Xavier-etal-IJOC-DATA).  The following files were downloaded to `/data_raw/Xavier/`

```{r echo=FALSE}
list.files(here::here("data_raw", "Xavier"), pattern = ".nc$")
```


# Data processing overview

After .nc files were downloaded they were roughly cropped to an area containing the Biological Dynamics of Forest Fragments Project (BDFFP).

- xmin: -65ºE
- xmax: -50ºE
- ymin: -5ºN
- ymax: 0ºN

Then values for each variable were aggregated spatially by taking the mean value for a 200km radius circle around the estimated center of BDFFP (-59.833ºE, -2.41ºN).  This aggregation was "exact", meaning it used a weighted mean, weighting raster pixels that were only partially contained in the circle appropriately.

# Column names and units

The finished data are found in the `data/` directory with one .csv file per model (AKA `source_id`) with the following structure:

```{r include=FALSE}
meta <- structure(
  list(
    colname = c(
      "dir",
      "source_id",
      "experiment_id",
      "time",
      "hfls",
      "hfss",
      "pr",
      "tas",
      "tasmin",
      "tasmax"
    ),
    longname = c(
      "directory",
      "source ID",
      "experiment ID",
      "time",
      "Surface Upward Latent Heat Flux",
      "Surface Upward Sensible Heat Flux",
      "Precipitation",
      "Near-Surface Air Temperature",
      "Daily Maximum Near-Surface Air Temperature",
      "Daily Minimum Near-Surface Air Temperature"
    ), 
    type = c(
      "character",
      "character",
      "character",
      "POSIXct",
      "numeric",
      "numeric",
      "numeric",
      "numeric",
      "numeric",
      "numeric"
    ),
    units = c(NA, NA, NA, NA, "W m-2",
              "W m-2", "kg m-2 s-1", "K", "K", "K")
  ),
  row.names = c(NA,-10L),
  class = "data.frame"
)
```


```{r echo=FALSE}
kable(meta)
```

# Reproducibility

To reproduce download and analysis run the scripts in `R/` sequentially.  Please note that this will take a very long time (possibly 24+ hrs) because of the purposefully slow download script (to avoid bans due to rate limits). Downloads from Copernicus use a modified Python script to access their API loaded into R using the `reticulate` package. To access the Compernicus API, you'll need to go through the steps described here: https://cds.climate.copernicus.eu/api-how-to

# QA/QC

See [data validation](https://brunalab.github.io/bdffp-cmip6/) for reports and plots of the data.  See also the errata for CMIP6 to check for known issues with particular models, scenarios, and variables (https://errata.es-doc.org/static/index.html)

# Citation

If you wish to use or adapt the code or data sets in this repository, please cite as:

>Eric R Scott. (2022). BrunaLab/bdffp-cmip6: Downloading and wrangling CMIP6 model output for the Biological Dynamics of Forest Fragments Project region. Zenodo. https://doi.org/10.5281/zenodo.6558263
