---
title: "Plot and validate CMIP6 downloads"
knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pointblank)
library(tidyverse)
library(here)
library(patchwork)
library(lubridate)
library(tsibble)
library(units)
library(ggforce) #to get ggplot to work with `units`
library(gt)
library(SPEI)
library(glue)
source(here::here("R", "functions.R"))
theme_set(theme_bw())
# Use `validate_rmd()` here to set options for the
# pointblank validation workflow within R Markdown documents
```

```{r load_index, message=FALSE, warning=FALSE, include=FALSE}
idx <- read_csv(here("metadata", "cmip6_index.csv"))
```

# Check that all models were downloaded

Check that all the directories for the .nc files got made

```{r validate=TRUE}
source_dl <- dir(here("data_raw", "CMIP6"))
source_id <- idx$source_id %>% unique() %>% str_to_lower() %>% str_replace_all("-", "_")
stop_if_not(!any(!source_id %in% source_dl))
```

Check that all the corresponding .csv files exist

```{r validate=TRUE}
csvs <- list.files(here('data'))
stop_if_not(!any(!paste0(source_id, "_data.csv") %in% csvs))
```



```{r read_files, message=FALSE, warning=FALSE, include=FALSE}
#read in all files for CMIP6
df_list <-
  map(list.files(here('data'),
                 pattern = "data\\.csv$",
                 full.names = TRUE),
      read_csv) %>%
  set_names(map(., ~.x$source_id[1]))

# observed data
xa_raw <- read_csv(here("data", "xavier_aggregated.csv"))
xa <- 
  xa_raw %>%
  mutate(date = paste(yearmonth, 1) %>% ymd(),
         month = month(date, label = TRUE)) %>% 
  group_by(month) %>%
  summarize(across(c(pr, starts_with("tas")), mean)) %>%
  mutate(pr = set_units(pr, "mm/month"),
         across(starts_with("tas"), ~set_units(.x, "degC")))
```

```{r filter_time, include=FALSE}
#filter so timespan is consistent
df_list <- 
  df_list %>% map(~{
    .x %>% 
      filter((
        experiment_id == "historical" &
          between(
            time,
            ymd_hms("1850-01-01 12:00:00"),
            ymd_hms("2014-12-31 12:00:00")
          )
      ) |
        (
          experiment_id != "historical" &
            between(
              time,
              ymd_hms("2015-01-01 12:00:00"),
              ymd_hms("2100-12-31 12:00:00")
            )
        ))
    
  })
```

```{r convert, include=FALSE}
df_list <- map(df_list, convert_units)
```


# Ensure that all models are complete

For this analysis, I only want to use models with `pr`, `tas`, `hfss`, and `hfls` variables in all 5 scenarios (historical, ssp126, ssp245, ssp370, and ssp585)

```{r include=FALSE}
df_list <- 
  df_list %>%
  keep(~all(c("pr", "tas", "hfss", "hfls") %in% names(.x))) %>% 
  keep(~all(c("historical", "ssp126", "ssp245", "ssp370", "ssp585") %in% .x$experiment_id))
```

# Calculations

Perform necessary calculations to compare PET and SPEI among models and between models and observed.

For PET, I'm using the "energy-only" method proposed by Milly and Dune (2016) eq. 8:

$$
PET = 0.8(R_n - G)
$$

Except that in their notes, they estimate $R_n -G$ as `hfls` + `hfss` after converting to units of mm/day using the latent heat of vaporazation of water, given by their eq. 2:

$$
L_v(T) = 2.501 - 0.002361T
$$
in MJ/kg

```{r pet, include=FALSE}
df_list <- 
  df_list %>%
  map(~ .x %>% mutate(pet = pet_energy_only(hfls, hfss, tas))) %>% 
  #convert PET to mm/month
  map(~.x %>%
        mutate(pet = pet * set_units(days_in_month(time), "d/month"),
               cb = pr - pet)
  )
```

For the observed data and the CMIP6 data from the same period, I calculate 3-month SPEI using precipitation and PET.

```{r spei, include=FALSE}
df_list <- 
  df_list %>%
  map(calc_spei)

#calc spei for observed
xa_spei <-
  xa_raw %>% 
  mutate(cb = pr - eto,
         date = ymd(paste(yearmonth, "15"))) %>% 
  mutate(spei = as.numeric(
    spei(
      ts(cb, freq = 12, start = c(year(min(date)), month(min(date)))),
      scale = 3,
      ref.start = c(1980, 4),
      ref.end = c(2014, 12)
    )$fitted
  ))
```


# Comparison to Observed Historical

Because SPEI, the variable of interest, is standardized, we looked for models that captured seasonality of precipitation rather than focusing on how well they estimate the exact amounts of precipitation. We calculated mean precipitation for January through December and then calculated a correlation coefficient between the 12 monthly means from each CMIP model and the 12 observed means.  We eliminated models with correlation coefficients less than 0.6 for precipitation.  Additionally, SPEI takes evapotranspiration into account, specifically through `hfls` and `hfss`.  Because these variables were not available for observed data, we used temperature but with a less stringent cutoff of correlations greater than 0.4.  Additionally we calculated SPEI for each model and counted the number of droughts (SPEI < -1) in each month.  However, SPEI was not used to determine which models remained in our ensemble because we did not have strong expectations that accuracy of past SPEI or drought frequency is a good measure of GCM model skill in predicting future SPEI.


```{r message=FALSE, warning=FALSE, include=FALSE}
# get monthly means of precip and temperature
season_df <- 
  df_list %>% 
  map_dfr(~{.x %>% 
    filter(time >= ymd("1980-01-01") &
             time <= ymd("2015-09-30")) %>% 
      mutate(month = month(time), .after=time) %>% 
      group_by(month) %>% 
      summarize(across(c(pr, starts_with("tas")), mean,
                       .names = "mean_{.col}")) 
  }, .id = "source_id")

# get monthly means for observed
xa2 <- xa %>% 
  rename_with(~paste0("mean_", .), .cols = -month) %>% 
  add_column(source_id = "observed") %>% 
  mutate(month = as.numeric(month))
season_df <- bind_rows(xa2, season_df) %>% 
  #move "observed" to the top by making it the first level
  mutate(source_id = fct_inorder(source_id))
```

```{r}
cmip_drought <- 
  df_list %>%
  map_dfr(~.x %>%
            filter(time >= ymd("1980-01-01") &
                     time <= ymd("2015-09-30")) %>%
            calc_drought_duration, .id = "source_id")

cmip_drought <-
  bind_rows(
    calc_drought_duration(xa_spei) %>% add_column(source_id = "observed"),
    cmip_drought
  ) %>% 
  select(source_id, mean_n_mon, sd_n_mon)
```


```{r mini_plots, include=FALSE}
# Create mini plots of monthly average precip and temperature for table

# season_df$mean_pr %>% range()
# season_df$mean_tasmin %>% min()
# season_df$mean_tasmax %>% max()

season_plots <-
  season_df %>%  
  group_by(source_id) %>%
  nest() %>% 
  mutate(
    pr_season = map(data, ~{
      ggplot(.x, aes(x = month, y = mean_pr)) +
        geom_col(fill = "blue") +
        scale_x_continuous("month", breaks = 1:12,
                           labels = ~month(.x, label = FALSE)) +
        scale_y_unit("pr", limits = c(0, 456)) +
        labs(y = "mm/mon") +
        theme(text = element_text(size = 45))
    }),
    tas_season = map(data, ~{
      ggplot(.x, aes(x = month)) +
        geom_ribbon(aes(ymin = mean_tasmin, ymax = mean_tasmax),
                    fill = "red",
                    alpha = 0.5) +
        geom_line(aes(y = mean_tas), color = "red") +
        scale_x_continuous("month",
                           breaks = 1:12,
                           labels = ~ month(.x, label = FALSE)) +
        scale_y_unit("tas", limits = c(19, 43)) +
        labs(y = "ºC") +
        theme(text = element_text(size = 45))
    })) %>% 
      dplyr::select(-data)
```

```{r}
#calculate number of droughts per month and mean drought duration

# number droughts per month
drought_df <- 
  df_list %>%
  map_dfr( ~{ 
    .x %>% filter(time >= ymd("1980-01-01") &
                    time <= ymd("2015-09-30")) %>% 
      mutate(drought = spei < -1) %>% 
      filter(!is.na(drought)) %>% 
      mutate(month = month(time)) %>% 
      group_by(month) %>% 
      summarize(n_droughts = sum(drought))
  }, .id = "source_id")

xa_drought_seasonality <-
  xa_spei %>% 
  mutate(drought = spei < -1) %>% 
  filter(!is.na(drought)) %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>% 
  summarize(n_droughts = sum(drought)) %>% 
  add_column(source_id = "observed")

drought_df <- bind_rows(drought_df, xa_drought_seasonality)

```

```{r drought-plots}
#create mini-histograms of numbers of droughts

drought_plots <-
  drought_df %>%  
  group_by(source_id) %>%
  nest() %>% 
  mutate(
    drought_season = map(data, ~{
      ggplot(.x, aes(x = month, y = n_droughts)) +
        geom_col(fill = "orange") +
        scale_x_continuous("month", breaks = 1:12,
                           labels = ~month(.x, label = FALSE)) +
        labs(y = "# SPEI < -1") +
        theme(text = element_text(size = 45))
    })) %>% 
      dplyr::select(-data)
# drought_plots$drought_season[[1]]
```


```{r table_data, include=FALSE}
table_data <- 
  season_df %>%
  group_by(source_id) %>%
  summarize(
    pr_cor = cor(mean_pr, xa$pr, method = "pearson"),
    tas_cor = cor(mean_tas, xa$tas, method = "pearson")
  ) %>%
  full_join(cmip_drought, by = "source_id") %>% 
  #format column for mean ± sd
  mutate(duration = glue("{round(mean_n_mon, 1)}±{round(sd_n_mon, 1)}")) %>% 
  dplyr::select(-mean_n_mon, -sd_n_mon) %>% 
  arrange(
    min_rank(desc(pr_cor)) +
    min_rank(desc(tas_cor))
    ) %>% 
  full_join(season_plots) %>% 
  full_join(drought_plots)
```



```{r gt_rough, include=FALSE}
table_rough <-
  table_data %>% 
  #make columns for plots to go eventually
  mutate(pr_plot = NA,
         tas_plot = NA,
         drought_plot = NA) %>%
  select(source_id, pr_cor, pr_plot, tas_cor, tas_plot, drought_duration = duration, drought_plot) %>% 
  #make into gt
  gt() %>%
  #place plots
  text_transform(
    locations = cells_body(pr_plot),
    fn = function(x) {
      map(table_data$pr_season,
          ggplot_image,
          height = px(80),
          aspect_ratio = 1.5)
    }
  ) %>% 
  text_transform(
    locations = cells_body(tas_plot),
    fn = function(x) {
      map(table_data$tas_season,
          ggplot_image,
          height = px(80),
          aspect_ratio = 1.5)
    }
  ) %>% 
  text_transform(
    locations = cells_body(drought_plot),
    fn = function(x) {
      map(table_data$drought_season,
          ggplot_image,
          height = px(80),
          aspect_ratio = 1.5)
    }
  ) %>% 
  # round correlations
  fmt_number(ends_with("_cor"))
# table_rough
```

```{r gt_format, echo=FALSE, message=FALSE, warning=FALSE}
table_rough %>% 
  # highlight poor correlations
  tab_style(
    style = list(
      cell_text(color = "red")
    ),
    locations = cells_body(
      columns = pr_cor,
      rows = pr_cor < 0.6
    )
  ) %>% 
    tab_style(
    style = list(
      cell_text(color = "red")
    ),
    locations = cells_body(
      columns = tas_cor,
      rows = tas_cor < 0.4
    )
  )  %>% 
  tab_spanner("Seasonality (monthly means)",
              c(starts_with("pr_"), starts_with("tas_")),
              id = "season") %>%
  tab_spanner("Droughts (SPEI < -1)", c(drought_duration, drought_plot), id = "droughts") %>%
  cols_label(
    source_id = "Source",
    pr_cor = "pr cor",
    tas_cor = "tas cor",
    pr_plot = "precipitation",
    tas_plot = "temperature",
    drought_duration = "mean ± SD duration (months)",
    drought_plot = "drought seasonality" 
  ) %>% 
  tab_footnote(
    md("Observed data from [Xavier et al. (2016)](https://doi.org/10.1002/joc.4518)"),
    cells_body(
      columns = source_id,
      rows = source_id == "observed"
    )
  ) %>% 
  tab_footnote("Red numbers highlight correlations (Pearson's r) < 0.6 for precipitation and < 0.4 for mean temperature.",
               cells_column_spanners("season")) %>% 
  tab_header(title = "Comparison of observed data to CMIP6 'historical' output",
             subtitle = "Data only from 1980 to 2015 to match observed.")
```


# CMIP model details

Below are plots of all data downloaded from each CMIP6 source.

```{r plot_fun, include=FALSE}
# Function to plot data
plot_all <- function(data) {
  p_list <- map(c("tas", "pr", "hfls", "hfss", "pet", "spei"),
      ~ggplot(data,
              aes_string(x = "time", y = .x, color = "experiment_id")) +
        geom_line(alpha = 0.5) + 
        geom_smooth(se = FALSE)
  ) 
  patchwork::wrap_plots(p_list, guides = "collect") &
    plot_annotation(title = data$source_id[1])
}
```


```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
for (i in seq_len(length(df_list))) {
  cat('\n')
  cat("## ", names(df_list[i]), "\n")
  p <- plot_all(df_list[[i]])
  print(p)
  cat('\n')
}
```




